Summary of the things that I learned

### 2024-10-22 100-introduction.tex
Process feedback about introduction, conclusion, bibliography and citations

Later, all kinds of different flavours and extensions of the $ \lambda $-calculus were put forth, with colorful names like `simply typed $ \lambda $-calculus', `System T' and `PCF'. Even though the $ \lambda $-calculus was originally a very theoretical tool, it was also the inspiration for function programming languages, and traces of it can be seen in imperative programming languages, where unnamed functions are commonly called `lambda expressions' \autocite{java-lambdas} and are sometimes even written like \texttt{lambda x y : (x - y) * (x + y)} \autocite{python-expressions}.

Even so, the theoretical study of the $ \lambda $-calculus and its extensions continues to this day. For example, in 2017\footnote{Note that the paper has been around since 2012, when it was first published as a preprint on arXiv.}, a paper by Martin Hyland was published with the title `classical lambda calculus in modern dress' \autocite{Hyland}. In this paper, Hyland approaches the $ \lambda $-calculus from the viewpoint of universal algebra, using algebraic theories, and more generally category theory, to study it. This way, he obtains two new proofs for old theorems. The paper also contains a new theorem that shows that two different ways to study the $ \lambda $-calculus using universal algebra are equivalent.

Now, in the last century, mathematics has changed a lot. Of course, new theorems have been proved, new conjectures have been made and entire new areas of mathematics have come into existence. However, as in many professions, the arrival of the computer has affected the way that work is done in mathematics. All kinds of tools have been created that aid mathematicians in their job. Some of these tools help with quick calculations, for formulating or disproving new conjectures. Other tools help in structurally verifying ideas. There even have been some `proofs by computer', which consist of a proof on paper that a theorem can be reduced to a finite, but very large, computation, and a computer program that then executes this computation. For example, the first proofs of the four color theorem \autocite{four-color-theorem} and the Kepler conjecture \autocite{Kepler-conjecture} were done this way.

However, because the computation part of such a proof involves a lot of code, and computer code tends to contain bugs, accepting a proof by computer involves a certain amount of trust. Therefore, both of the theorems mentioned have subsequently also been proved using `computer proof assistants' \autocite{formalized-four-color-theorem}\autocite{formalized-Kepler-conjecture}. Such proof assistants, like `rocq', `lean' and `Agda', are computer programs with only a very small `trusted codebase'\footnote{The trusted codebase is the part of the codebase of which we have to trust that it is correct. The rest of the codebase is checked using this trusted core, so if the trusted core is correct, the rest is too.}, that can verify mathematical reasoning. In this way, if we trust the small core of such a proof assistant, and the proof assistant says that a proof is correct, then we can trust that indeed, the proof is correct.

@@

Now, much of contemporary mathematics is built on the foundation of set theory, which we will often refer to as \iindex{classical mathematics}. Usually, `set theory' refers to ZFC: Zermelo Fraenkel set theory with the axiom of choice, developed in the early 1900s \autocite{Zermelo-set-theory}. On the other hand, most of the formalization in proof assistants is built on the foundation of type theory. This is because computers can reason better with types than with sets, as evidenced by the many typed programming languages that are around nowadays. However, type theory has been around longer than computers have: it was initially developed by Bertrand Russell in the early 1900s (\autocite{Russell-type-theory}, Appendix B), to create a foundation of mathematics that avoided Russell's paradox\footnote{Russell's paradox is the question ``Consider the set of all sets that do not contain themself. Does this set contain itself?'' Answering this question with either `yes' or `no' leads to a contradiction.}.

@@

In this thesis, we study the paper `classical lambda calculus in modern dress' through the lens of univalent foundations, work out the details of the proofs, and formalize part of the paper and its preliminaries. The major contributions are:
\begin{itemize}
  \item More detailed versions of Hyland's definitions and proofs, complemented with some examples (Chapters \ref{ch:algebraic-structures} and \ref{ch:the-paper}). Also, the work of Hyland's predecessors that he expands upon (Chapter \ref{ch:previous-work}), and most of the category theoretical preliminaries needed to understand the paper (Chapter \ref{ch:category-theory}), to make this thesis reasonably self-contained. All this makes Hyland's paper more accessible to computer scientists.
  \item The translation of Hyland's paper and the work of his predecessors to univalent foundations (Chapter \ref{ch:the-paper}). This shows that his work can be translated to univalent foundations, and contributes to the knowledge about translating classical mathematics to univalent foundations in general.
  \item A formalization of part of the paper in rocq (Chapter \ref{ch:the-formalization}). This contributes to the knowledge about formalization in general.
  \item A new proof for the fundamental theorem of the $ \lambda $-calculus (Section \ref{sec:elementary-fundamental-theorem}). This version is easier to read and verify, because it is more elementary and uses less category theory than Hyland's proof.
  \item An analysis of the behaviour of the Karoubi envelope in univalent foundations in general (Section \ref{sec:karoubi-envelope}), and in the specific case of Paul Taylor's and Martin Hyland's work (Remark \ref{rem:difference-Taylor-Hyland}). Here we see an example of two definitions that are equivalent in classical mathematics, but become different in univalent foundations. It is interesting to see the differences between the behaviour of the two definitions in univalent foundations.
\end{itemize}

>>>>

Later, all kinds of different flavours and extensions of the $ \lambda $-calculus were put forth, with colorful names like `simply typed $ \lambda $-calculus', `System T' and `PCF'. Even though the $ \lambda $-calculus was originally a very theoretical tool, it was also the inspiration for functional programming languages, and traces of it can be seen in imperative programming languages, where unnamed functions are commonly called `lambda expressions' \autocite{java-lambdas} and are sometimes even written like \texttt{lambda x y : (x - y) * (x + y)} \autocite{python-expressions}.

Even so, the theoretical study of the $ \lambda $-calculus and its extensions continues to this day. For example, in 2017\footnote{Note that the paper has been around since 2012, when it was first published as a preprint on arXiv.}, a paper by Martin Hyland was published with the title `Classical lambda calculus in modern dress' \autocite{Hyland}. In this paper, Hyland approaches the $ \lambda $-calculus from the viewpoint of universal algebra, using algebraic theories, and more generally category theory, to study it. This way, he obtains two new proofs for old theorems. The paper also contains a new theorem that shows that two different ways to study the $ \lambda $-calculus using universal algebra are equivalent.

Now, in the last century, mathematics has changed a lot. Of course, new theorems have been proved, new conjectures have been made and entire new areas of mathematics have come into existence. However, as in many professions, the arrival of the computer has affected the way that work is done in mathematics. All kinds of tools have been created that aid mathematicians in their job. Some of these tools help with quick calculations for formulating or disproving new conjectures. Other tools help in structurally verifying ideas. There even have been some `proofs by computer', which consist of a proof on paper that a theorem can be reduced to a finite, but very large, computation, and a computer program that then executes this computation. For example, the first proofs of the four color theorem \autocite{four-color-theorem} and the Kepler conjecture \autocite{Kepler-conjecture} were done this way.

However, because the computation part of such a proof involves a lot of code, and computer code tends to contain bugs, accepting a proof by computer involves a certain amount of trust. Therefore, both of the theorems mentioned have subsequently also been proved using `computer proof assistants' \autocite{formalized-four-color-theorem, formalized-Kepler-conjecture}. Such proof assistants, like `rocq', `lean' and `Agda', are computer programs with only a very small `trusted codebase'\footnote{The trusted codebase is the part of the codebase of which we have to trust that it is correct. The rest of the codebase is checked using this trusted core, so if the trusted core is correct, the rest is too.}, that can verify mathematical reasoning. In this way, if we trust the small core of such a proof assistant, and the proof assistant says that a proof is correct, then we can trust that indeed, the proof is correct.

@@

Now, much of contemporary mathematics is built on the foundation of set theory, which we will often refer to as \iindex{classical mathematics}. Usually, `set theory' refers to ZFC: Zermelo Fraenkel set theory with the axiom of choice, developed in the early 1900s \autocite{Zermelo-set-theory}. On the other hand, most of the formalization in proof assistants is built on the foundation of type theory. This is because computers can reason better with types than with sets, as evidenced by the many typed programming languages that are around nowadays. However, type theory has been around longer than computers have: it was initially developed by Bertrand Russell in the early 1900s \autocite[][Appendix B]{Russell-type-theory}, to create a foundation of mathematics that avoided Russell's paradox\footnote{Russell's paradox is the question ``Consider the set of all sets that do not contain themself. Does this set contain itself?'' Answering this question with either `yes' or `no' leads to a contradiction.}.

@@

In this thesis, we study the paper `Classical lambda calculus in modern dress' through the lens of univalent foundations, work out the details of the proofs, and formalize part of the paper and its preliminaries. The major contributions are:
\begin{itemize}
  \item More detailed versions of Hyland's definitions and proofs, complemented with some examples (Chapters \ref{ch:algebraic-structures} and \ref{ch:the-paper}). Also, the work of Dana Scott and Paul Taylor that Hyland expands upon (Chapter \ref{ch:previous-work}), and most of the category theoretical preliminaries needed to understand the paper (Chapter \ref{ch:category-theory}), to make this thesis reasonably self-contained. All this makes Hyland's paper more accessible to computer scientists.
  \item The translation of Hyland's paper and the work of his predecessors to univalent foundations (Chapter \ref{ch:the-paper}). This shows that his work can be translated to univalent foundations, and contributes to the knowledge about translating classical mathematics to univalent foundations in general.
  \item A formalization of part of the paper in rocq (Chapter \ref{ch:the-formalization}). This contributes to the knowledge about formalization in general.
  \item A new proof for the fundamental theorem of the $ \lambda $-calculus (Section \ref{sec:elementary-fundamental-theorem}). This version is easier to read and verify, because it is more elementary and uses less category theory than Hyland's proof.
  \item An analysis of the behaviour of the Karoubi envelope in univalent foundations in general (Section \ref{sec:karoubi-envelope}), and in the specific case of Paul Taylor's and Martin Hyland's work (Remark \ref{rem:difference-Taylor-Hyland}). Here we see an example of two definitions that are equivalent in classical mathematics, but become different in univalent foundations. It is interesting to see the differences between the behaviour of the two definitions in univalent foundations.
\end{itemize}

### 2024-10-22 200-category-theory.tex
Process feedback about introduction, conclusion, bibliography and citations

An alternative characterization (see \autocite{MacLane}, chapter IV.1, Theorem 2) of an adjunction $ L \dashv R $ is as a natural bijection
\[ \varphi: \D(L(X), Y) \xrightarrow{\sim} \C(X, R(Y)). \]
Naturality means that for all $ f: \C(X^\prime, X) $, $ g: \D(Y, Y^\prime) $ and $ h: \D(L(X), Y) $,
\[ \varphi(L(f) \cdot h \cdot g) = f \cdot \varphi(h) \cdot R(g). \]

@@

We can embed $ \C $ fully faithfully into $ P \C $ as follows (see \autocite{Kashiwara}, Section 1.4):
\begin{definition}\label{def:Yoneda-embedding}
  The \iindex{Yoneda embedding} $ \yo : \C \hookrightarrow P \C $ is given on objects by $ \yo(Y) = \C(-, Y) $:
  \[ \yo(Y)(X) = \C(X, Y) \quad \text{and} \quad \yo(Y)(f)(g) = f \cdot g \]
  for $ X: \C $, $ f: \C(X, X^\prime) $ and $ g: \C(X^\prime, Y) $. It sends a morphism $ f: \C(Y, Y^\prime) $ to the natural transformation $ \yo(f): \C(-, Y) \Rightarrow \C(-, Y^\prime) $ given by
  \[ \yo(f)(X)(g) = g \cdot f \]
  for $ X: \C $ and $ g: \C(X, Y) $.
\end{definition}

Now, this embedding has a couple of properties:
\begin{lemma}
  For any $ Y: \C $ and $ F : P \C $, we have an equivalence $ P \C(\yo(Y), F) \simeq F(Y) $, and this equivalence is natural in $ Y $ and $ F $.
\end{lemma}
\begin{proof}
  It sends a natural transformation $ a: P \C(\yo(Y), F) $ to the element $ a_Y(\id Y) $. Conversely, it sends an element $ X : P(Y) $ to the natural transformation $ a $ given by $ a_X(f) = P(f)(X) $ for all $ f: \yo(Y)(X) = \C(X, Y) $. For more details, see \autocite{Kashiwara}, Proposition 1.4.3.
\end{proof}

@@

\begin{remark}
  Also, for any small category $ \C $ (small means that the collection of objects is a type in our type universe. For example: for a fixed universe of types, $ \TYPE $ or $ \SET $ are not small categories, the universe of types is not contained in itself), the presheaf category $ P \C $ has exponential objects, given by
  \[ (F^G)(X) = P\C(\yo(X) \times G, F), \]
  the natural transformations from the product functor of the Yoneda embedding of $ X $ and $ G $, to $ F $ (see \autocite{MacLane-Moerdijk}, Section I.6, Proposition 1).
\end{remark}

@@

\begin{lemma}[\coqident{AlgebraicTheories.FundamentalTheorem.SurjectivePrecomposition}{pre_comp_split_essentially_surjective}]\label{lem:lan-precomp-iso}
  If $ F : \A \to \B $ is a fully faithful functor, and $ \C $ is a category with colimits, $ \eta: \id{[\A, \C]} \Rightarrow \Lan F {} \bullet F_* $ is a natural isomorphism.
\end{lemma}
\begin{proof}
  To show that $ \eta $ is a natural isomorphism, we have to show that $ \eta_G(Y): G(Y) \Rightarrow \Lan F G(F(Y)) $ is an isomorphism for all $ G: [\A, \C] $ and $ Y: \A $. Since a left adjoint is unique up to natural isomorphism (\autocite{CT4P}, Exercise 153), we can assume that $ \Lan F G(F(Y)) $ is given by
  \[ \text{colim} ((F \downarrow F(Y)) \to \A \xrightarrow G \C). \]
  Now, the diagram for this colimit consists of $ G(X) $ for each arrow $ f: \B(F(X), F(Y)) $. Since $ F $ is fully faithful, we have $ f = F(\overline f) $ for some $ \overline f: \A(X, Y) $. If we now take the arrows $ G(\overline f): \C(G(X), G(Y)) $, the universal property of the colimit gives an arrow
  \[ \varphi: \C(\Lan F G(F(Y)), G(Y)) \]
  which constitutes an inverse to $ \eta_G(Y) $. The proof of this revolves around properties of the colimit and its (induced) morphisms.
\end{proof}

>>>>

An alternative characterization \autocite[][Chapter IV.1, Theorem 2]{MacLane} of an adjunction $ L \dashv R $ is as a natural bijection
\[ \varphi: \D(L(X), Y) \xrightarrow{\sim} \C(X, R(Y)). \]
Naturality means that for all $ f: \C(X^\prime, X) $, $ g: \D(Y, Y^\prime) $ and $ h: \D(L(X), Y) $,
\[ \varphi(L(f) \cdot h \cdot g) = f \cdot \varphi(h) \cdot R(g). \]

@@

We can embed $ \C $ fully faithfully into $ P \C $ as follows \autocite[][Section 1.4]{Kashiwara}:
\begin{definition}\label{def:Yoneda-embedding}
  The \iindex{Yoneda embedding} $ \yo : \C \hookrightarrow P \C $ is given on objects by $ \yo(Y) = \C(-, Y) $:
  \[ \yo(Y)(X) = \C(X, Y) \quad \text{and} \quad \yo(Y)(f)(g) = f \cdot g \]
  for $ X: \C $, $ f: \C(X, X^\prime) $ and $ g: \C(X^\prime, Y) $. It sends a morphism $ f: \C(Y, Y^\prime) $ to the natural transformation $ \yo(f): \C(-, Y) \Rightarrow \C(-, Y^\prime) $ given by
  \[ \yo(f)(X)(g) = g \cdot f \]
  for $ X: \C $ and $ g: \C(X, Y) $.
\end{definition}

Now, this embedding has a couple of properties:
\begin{lemma}
  For any $ Y: \C $ and $ F : P \C $, we have an equivalence $ P \C(\yo(Y), F) \simeq F(Y) $, and this equivalence is natural in $ Y $ and $ F $.
\end{lemma}
\begin{proof}
  It sends a natural transformation $ a: P \C(\yo(Y), F) $ to the element $ a_Y(\id Y) $. Conversely, it sends an element $ X : P(Y) $ to the natural transformation $ a $ given by $ a_X(f) = P(f)(X) $ for all $ f: \yo(Y)(X) = \C(X, Y) $. For more details, see \autocite[][Proposition 1.4.3]{Kashiwara}.
\end{proof}

@@

\begin{remark}
  Also, for any small category $ \C $ (small means that the collection of objects is a type in our type universe. For example: for a fixed universe of types, $ \TYPE $ or $ \SET $ are not small categories, the universe of types is not contained in itself), the presheaf category $ P \C $ has exponential objects, given by
  \[ (F^G)(X) = P\C(\yo(X) \times G, F), \]
  the natural transformations from the product functor of the Yoneda embedding of $ X $ and $ G $, to $ F $ \autocite[][Section I.6, Proposition 1]{MacLane-Moerdijk}.
\end{remark}

@@

\begin{lemma}[\coqident{AlgebraicTheories.FundamentalTheorem.SurjectivePrecomposition}{pre_comp_split_essentially_surjective}]\label{lem:lan-precomp-iso}
  If $ F : \A \to \B $ is a fully faithful functor, and $ \C $ is a category with colimits, $ \eta: \id{[\A, \C]} \Rightarrow \Lan F {} \bullet F_* $ is a natural isomorphism.
\end{lemma}
\begin{proof}
  To show that $ \eta $ is a natural isomorphism, we have to show that $ \eta_G(Y): G(Y) \Rightarrow \Lan F G(F(Y)) $ is an isomorphism for all $ G: [\A, \C] $ and $ Y: \A $. Since a left adjoint is unique up to natural isomorphism \autocite[][Exercise 153]{CT4P}, we can assume that $ \Lan F G(F(Y)) $ is given by
  \[ \text{colim} ((F \downarrow F(Y)) \to \A \xrightarrow G \C). \]
  Now, the diagram for this colimit consists of $ G(X) $ for each arrow $ f: \B(F(X), F(Y)) $. Since $ F $ is fully faithful, we have $ f = F(\overline f) $ for some $ \overline f: \A(X, Y) $. If we now take the arrows $ G(\overline f): \C(G(X), G(Y)) $, the universal property of the colimit gives an arrow
  \[ \varphi: \C(\Lan F G(F(Y)), G(Y)) \]
  which constitutes an inverse to $ \eta_G(Y) $. The proof of this revolves around properties of the colimit and its (induced) morphisms.
\end{proof}

### 2024-10-22 300-univalent-foundations.tex
Process feedback about introduction, conclusion, bibliography and citations

So if we have types and elements, are types also elements of some type? This is a tricky question, because if we say that there exists a type $ \TYPE $ which contains all types (and therefore, itself), we introduce inconsistency in our type system (see \autocite{girard}). This is just like having a `set of all sets', which results in problematic questions like ``does its subset, containing only the sets which do not contain itself, contain itself?'' This is usually solved by either stating that types do not themselves have a type, or by assuming the existence of \textit{type universes} $ U_1, U_2, \dots $, such that every type is an element of some $ U_i $. Note that universes are allowed to be inclusive or not: we may have $ U_n \subseteq U_{n + 1} $. In this thesis, we will not explicitly mention particular universes. We will just write $ \TYPE $ to denote some type universe (or its category of types and functions). This is called \textit{typical ambiguity}, meaning that every theorem holds for all (suitable) assignments of universes to these instances of $ \TYPE $.

@@

It is important to note that Martin-Löf type theory has \textit{identity types}: given a type $ T $ and elements $ x, y: T $, we have a type $ \mathrm{Id}_T(x, y) $ (note that this is a dependent type), which we will usually denote with $ x = y $. An element $ p: x = y $ is a proof that $ x $ is `equal' to $ y $. This type comes with an interesting induction principle called \iindex{path induction}: we can show any statement about paths $ p: x = y $ for generic $ x $ and $ y $, if we can show it about the `trivial' path $ \mathrm{refl}: x = x $ for generic $ x $. For example, symmetry of the equality boils down to a function $ \prod_{x y : T} x = y \to y = x $. We construct this using path induction with the function that sends $ \mathrm{refl}: x = x $ to itself. For more information, see \autocite{hottbook}, Section 1.12.1.

In set-theoretic mathematics, there is the concept of a `bijection' $ S \simeq T $ of sets (or an isomorphism in the category $ \SET $), which is often treated as an equivalence. It consists of functions $ f: S \to T $ and $ g: T \to S $ with $ f \cdot g = \id S $ and $ g \cdot f = \id T $. In type theory, we have a similar concept, which is called `equivalence' (of types) $ S \simeq T $. Since bijections are not well-behaved for types that are not sets (see Section \ref{sec:homotopy-props-sets}), because in those cases, a function $ f: S \to T $ can have multiple distinct inverses. Therefore, we define $ S \simeq T := \sum_{f: S \to T} \mathrm{isequiv}(f) $, for some predicate $ \mathrm{isequiv}: (S \to T) \to \TYPE $ (see \autocite{hottbook}, Equation 2.4.10). However, intuitively we can still think of these as bijections.

@@

\begin{definition}
  For a type $ T $, a type $ \Vert T \Vert $ exists (\autocite{hottbook}, Section 3.7) with the properties that for all $ t: T $, we have an element $ \vert t \vert : \Vert T \Vert $ and that $ \Vert T \Vert $ is a mere proposition. It has a recursion principle stating that for a mere proposition $ B $, a function $ f: T \to B $ induces a function $ \vert f \vert: \Vert T \Vert \to B $ that commutes with $ \vert \cdot \vert $. This object is called the \index{truncation!propositional}\textit{propositional truncation}.
\end{definition}
The propositional truncation forgets the details about a type, and only keeps the information about whether it is inhabited or not. The recursion principle means that if we are trying to prove a mere proposition based on some element $ \vert t \vert: \Vert T \Vert $, we can pretend that we do have a concrete element $ t: T $.

>>>>

So if we have types and elements, are types also elements of some type? This is a tricky question, because if we say that there exists a type $ \TYPE $ which contains all types (and therefore, itself), we introduce inconsistency in our type system \autocite{girard}. This is just like having a `set of all sets', which results in problematic questions like ``does its subset, containing only the sets which do not contain itself, contain itself?'' This is usually solved by either stating that types do not themselves have a type, or by assuming the existence of \textit{type universes} $ U_1, U_2, \dots $, such that every type is an element of some $ U_i $. Note that universes are allowed to be inclusive or not: we may have $ U_n \subseteq U_{n + 1} $. In this thesis, we will not explicitly mention particular universes. We will just write $ \TYPE $ to denote some type universe (or its category of types and functions). This is called \textit{typical ambiguity}, meaning that every theorem holds for all (suitable) assignments of universes to these instances of $ \TYPE $.

@@

It is important to note that Martin-Löf type theory has \textit{identity types}: given a type $ T $ and elements $ x, y: T $, we have a type $ \mathrm{Id}_T(x, y) $ (note that this is a dependent type), which we will usually denote with $ x = y $. An element $ p: x = y $ is a proof that $ x $ is `equal' to $ y $. This type comes with an interesting induction principle called \iindex{path induction}: we can show any statement about paths $ p: x = y $ for generic $ x $ and $ y $, if we can show it about the `trivial' path $ \mathrm{refl}: x = x $ for generic $ x $. For example, symmetry of the equality boils down to a function $ \prod_{x y : T} x = y \to y = x $. We construct this using path induction with the function that sends $ \mathrm{refl}: x = x $ to itself. For more information, see \autocite[][Section 1.12.1]{hottbook}.

In set-theoretic mathematics, there is the concept of a `bijection' $ S \simeq T $ of sets (or an isomorphism in the category $ \SET $), which is often treated as an equivalence. It consists of functions $ f: S \to T $ and $ g: T \to S $ with $ f \cdot g = \id S $ and $ g \cdot f = \id T $. In type theory, we have a similar concept, which is called `equivalence' (of types) $ S \simeq T $. Since bijections are not well-behaved for types that are not sets (see Section \ref{sec:homotopy-props-sets}), because in those cases, a function $ f: S \to T $ can have multiple distinct inverses. Therefore, we define $ S \simeq T := \sum_{f: S \to T} \mathrm{isequiv}(f) $, for some predicate $ \mathrm{isequiv}: (S \to T) \to \TYPE $ \autocite[][Equation 2.4.10]{hottbook}. However, intuitively we can still think of these as bijections.

@@

\begin{definition}
  For a type $ T $, a type $ \Vert T \Vert $ exists \autocite[][Section 3.7]{hottbook} with the properties that for all $ t: T $, we have an element $ \vert t \vert : \Vert T \Vert $ and that $ \Vert T \Vert $ is a mere proposition. It has a recursion principle stating that for a mere proposition $ B $, a function $ f: T \to B $ induces a function $ \vert f \vert: \Vert T \Vert \to B $ that commutes with $ \vert \cdot \vert $. This object is called the \index{truncation!propositional}\textit{propositional truncation}.
\end{definition}
The propositional truncation forgets the details about a type, and only keeps the information about whether it is inhabited or not. The recursion principle means that if we are trying to prove a mere proposition based on some element $ \vert t \vert: \Vert T \Vert $, we can pretend that we do have a concrete element $ t: T $.
